### 2020-12-17 01_introduction.tex
Updates from Overleaf

\IEEEPARstart{A}{utomation} tasks in industrial and service robotics, such as product packing or sorting, often require sets of objects to be arranged in
specific poses on a planar surface. Efficient and high-quality
single-arm solutions have been proposed for such
setups \cite{193}. The proliferation of robot arms, however, including dual-arm setups, implies that industrial settings can utilize multiple robots in the same workspace (Fig \ref{fig:two_arms}). This work explores a) the benefits of coordinated dual-arm rearrangement versus single-arm, b) the combinatorial challenges involved and c) computationally efficient, high-quality and scalable methods.

@@

A motivating point is that the coordinated use of multiple arms can
result in significant improvements in efficiency. This arises from the following argument.

% \vspace{-.1in}
\begin{lemma}\label{l:k-arm-lower-bound}
There are classes of tabletop rearrangement problems, where a $k$-arm
($k \ge 2$) solution can be arbitrarily better than the optimal
single-arm one.
% \vspace{-.3in}
\end{lemma}

@@

@@

% \vspace{-.1in}
\begin{lemma}\label{l:2-arm-no-worse}
For any rearrangement problem, the best $k$-arm ($k \ge 2$) solution
cannot be worse than an optimal single arm solution.
% \vspace{-.1in}
\end{lemma}

The above points motivate the development of scalable algorithmic
tools for such dual-arm rearrangement instances. This work considers
certain relaxations to achieve this objective. In particular,
monotone tabletop instances are considered, where the start and goal
object poses do not overlap.  Furthermore, the focus is on
synchronized execution of pick-and-place actions by the two arms. i.e., where the two arms simultaneously transfer (different) objects, or simultaneously move towards picking the next objects.
Theoretical arguments and experimental evaluation indicate that this does not significantly degrade solution
quality.

\begin{figure}[t]
% \vspace{-0.15in}
	\centering
	\includegraphics[width=0.48\textwidth]{figures/two_arms}
% 	\vspace{-0.1in}
	\caption{Example of dual-arm setups that can utilize
	algorithms proposed in this work.}
	\label{fig:two_arms}
% 	\vspace{-0.25in}
\end{figure}

The first contribution is the study of the combinatorial structure of
synchronous, monotone dual-arm rearrangement. Then, a mixed integer linear programming (\milp) model is
proposed that achieves optimal coordinated solutions in this domain.
The proposed efficient algorithm {\tt Tour\_Over\_Matching} (\algo) significantly improves in scalability. \algo\ first optimizes the
cost of object transfers and assigns objects to the two arms by
solving an optimal matching problem. It minimizes the cost of
moves from an object's goal to another object's start pose per
arm as a secondary objective by employing a {\tt TSP} solution. Most
of the computation time is spent on the many calls to a
lower-level motion planner that coordinates the two arms. A lazy
evaluation strategy is proposed, where motion plans are evaluated for candidate solution sequences. This results in significant computational improvement and
reduces the calls to the motion planner.  An analysis
studies the expected improvement in solution quality
versus the single arm case, as well as the expected cost overhead from a
synchronous solution.

Finally, experiments for i) a simple planar picker setting, and ii) two 7-DOF Kuka arms, demonstrate a nearly two-fold
improvement against the single arm case for the proposed approach in practice. The algorithm exhibits close to optimal solutions and good scalability. 
The lazy evaluation strategy significantly improves computation costs for both the optimal and proposed methods.
%The computational costs of both the optimal
%and the proposed method are significantly improved from the lazy
%evaluation strategy. 
% When the synchronous solutions are smoothed to
% asynchronous ones, there is not significance difference in overall
% path quality. \kiril{This statement is unclear. What is the benefit from performing this step?}

>>>>

\IEEEPARstart{A}{utomation} tasks in industrial and service robotics, such as product packing or sorting, often require sets of objects to be arranged in
specific poses on a planar surface. Efficient and high-quality
single-arm solutions have been proposed for such
setups \cite{193}. The proliferation of robot arms, however, including dual-arm setups, implies that industrial settings can utilize multiple robots in the same workspace (Fig~\ref{fig:two_arms}). This work explores a) the benefits of coordinated dual-arm rearrangement versus single-arm, b) the combinatorial challenges involved and c) computationally efficient, high-quality and scalable methods.

@@

A major motivation for this work is that the coordinated use of multiple arms can
result in significant improvements in efficiency. This arises from the following argument.

% \vspace{-.1in}
\begin{observation}\label{l:k-arm-lower-bound}
There are classes of tabletop rearrangement problems, where a $k$-arm
($k \ge 2$) solution can be arbitrarily better than the optimal
single-arm one.
% \vspace{-.3in}
\end{observation}

@@

\begin{figure}[t]
% \vspace{-0.15in}
	\centering
	\includegraphics[width=0.48\textwidth]{figures/two_arms}
% 	\vspace{-0.1in}
	\caption{Example of dual-arm setups that can utilize
	algorithms proposed in this work.}
	\label{fig:two_arms}
	\vspace{-0.25in}
\end{figure}

@@

% \vspace{-.1in}
\begin{observation}\label{l:2-arm-no-worse}
For any rearrangement problem, the best $k$-arm ($k \ge 2$) solution
cannot be worse than an optimal single arm solution.
% \vspace{-.1in}
\end{observation}

The above observations motivate the development of scalable algorithmic
tools for such dual-arm rearrangement instances. This work considers
certain relaxations to achieve this objective. In particular,
monotone tabletop instances are considered, where the start and goal
object poses do not overlap.  Furthermore, the focus is on
synchronized execution of pick-and-place actions by the two arms. i.e., where the two arms simultaneously transfer (different) objects, or simultaneously move towards picking the next objects.
Theoretical arguments and experimental evaluation indicate that this does not significantly degrade solution
quality.

The first contribution is the study of the combinatorial structure of
synchronous, monotone dual-arm rearrangement. Then, a mixed integer linear programming (\milp) model is
proposed that achieves optimal coordinated solutions in this domain.
The proposed efficient algorithm {\tt Tour\_Over\_Matching} (\algo) significantly improves in scalability. \algo\ first optimizes the
cost of object transfers and assigns objects to the two arms by
solving an optimal matching problem. It minimizes the cost of
moves from an object's goal to another object's start pose per
arm as a secondary objective by employing a {\tt TSP} solution. Most
of the computation time is spent on the many calls to an underlying motion planner that coordinates the two arms. A lazy
evaluation strategy is proposed, where motion plans are evaluated and validated for candidate solution sequences, as opposed to performing this expensive motion plan computation for the entire search space. This results in significant computational improvement and
reduces the calls to the motion planner.  An analysis
studies the expected improvement in solution quality
versus the single arm case, as well as the expected cost overhead from a
synchronous solution.

Finally, experiments for i) a simple planar picker setting, and ii) two 7-DOF Kuka arms, demonstrate a nearly two-fold
improvement against the single arm case for the proposed approach in practice. The algorithm exhibits close to optimal solutions and good scalability. 
The lazy evaluation strategy significantly improves computation costs for both the optimal and proposed methods.

\tase{
The current archival version is an extension of previous work \cite{shome2018dualarm}. In particular, several additions have been made to provide a more holistic study of the problem:
\begin{enumerate}
    % \item Additional explanation and visual support is provided for understanding the algorithmic steps, especially regarding Section \ref{sec:approximation}.
    \item A detailed illustration and expository text is provided for describing the algorithmic steps, especially regarding Section \ref{sec:approximation}.
	\item Further analysis has been performed to study the problem (Section \ref{sec:cost_bounds}). Specifically, the result describing the asymptotic cost bound for asynchronous operation has been extended to the $k$-arm case. Additionally, a more detailed proof is provided for the asymptotic cost bound estimate in the case of synchronized dual-arm operation.
% 	\item New data (Section \ref{sec:smoothing}) is provided that study the effect of post-processing solutions so as to minimize delays introduced due to the synchronization assumption. 
	\item The solutions from the benchmarks were post-processed so as to minimize delays introduced due to the synchronization assumption, and the data from this \textit{smoothing} operation is provided in Section \ref{sec:smoothing}.
	\item A benchmark (Section \ref{sec:reachability}) has been included to demonstrate the efficacy of the proposed approach in problem domains where initial or target poses of objects do not lie in the shared workspace.
\end{enumerate}
}

%The computational costs of both the optimal
%and the proposed method are significantly improved from the lazy
%evaluation strategy. 
% When the synchronous solutions are smoothed to
% asynchronous ones, there is not significance difference in overall
% path quality. \kiril{This statement is unclear. What is the benefit from performing this step?}

### 2020-12-17 02_background.tex
Updates from Overleaf

%Related efforts:
%
%\begin{itemize}
%\item multi-robot planning and coordination
%\item manipulation task planning in general
%\item rearrangement 
%\item combinatorial literature
%\end{itemize}
The current work dealing with dual-arm object rearrangement touches upon the challenging intersection of a variety of rich bodies of prior work. It is closely related to multi-robot planning and coordination where a challenge is the high dimensionality of the configuration space. %Early work has suggested to slightly reduce the~\cite{aronov1999motion} tried dimensionality reduction strategies. 
Optimal strategies were developed for simpler instances of the problem \cite{solovey2015motion}, although in general the problem is known to be computationally hard~\cite{solovey2016hardness}. Decentralized approaches~\cite{van2005prioritized} also used velocity tuning~\cite{leroy1999multiple}. General multi-robot planning tries to plan for multiple high-dimensional platforms \cite{wagner2012probabilistic,Gharbi:2009fu} using sampling-based techniques. 
Recent advances provide scalable~\cite{SoloveySH16:ijrr} and asymptotically optimal~\cite{Dobson:2017aa} sampling based frameworks. 
%Assembly planning \cite{halperin2000general,sundaram2001disassembly} is another related problem dealing with multi-body planning.

@@

%\textbf{The Traveling Salesman Problem}:
In this work we leverage a connection between dual-arm rearrangement and two combinatorial problems: (1) optimal matching~\cite{edmonds1965maximum} and (2) TSP. On the surface the problem seems closely related to multi-agent TSP.
%This can itself be tweaked with assumptions on same or different start and goal positions, symmetric or asymmetric metrics, directed or undirected graphs or posing the problem as an optimization
%A seminal paper~\cite{frederickson1976approximation} provides the formulation for k-TSP, with solutions that split a single tour.
%Prior work \cite{rathinam2006matroid} has formulated the problem of multi-start to multi-goal k-TSP as an optimization task. 
Prior work has formulated the k-TSP solution in terms of splitting a single tour~\cite{frederickson1976approximation} or as an optimization task~\cite{rathinam2006matroid}.
Some work~\cite{friggstad2013multiple} deals with asymmetric edge weights which are more relevant to the problems of our interest. 
%\textbf{Errand Scheduling}:
%If the object transfer is seen as an errand \cite{slavik1997errand} then a graph constructed with nodes being object-arm assignments and edges implying the order of coordinated execution. 
%There has been prior work \cite{garg2000polylogarithmic} that operates over the assignment graph(where pairwise assignments are nodes), an object's group consists of sets of nodes that include an object
The problem can be posed as an instance of multi vehicle pickup and delivery (PDP)~\cite{parragh2008survey}. 
%Most of the work has been motivated by ride-sharing applications and timed task scheduling, so there are a lot of variations that deal with time windows but this line of research provides some ILP formulations for the problem.
Prior work~\cite{coltin2014multi} has applied the PDP problem to robots, taking into account time windows and robot-robot transfers.
Some seminal work~\cite{lenstra1981complexity,savelsbergh1995general} has explored its complexity, and concedes to the hardness of the problem, while others studied cost bounds~\cite{TrePavFra13}, and proposed ILP formulations~\cite{savelsbergh1995general}. 
%The existing work that delves into the combinatorial challenges of the problem, do not account for the costs incurred due to coordination of the agents during task execution. 
%Pickup and Delivery problems under track contention \cite{caricato2003parallel}, makes an attempt to reason about the coordination of the robots over solutions obtained with Tabu Search.
Typically this line of work ignores coordination costs, though some methods~\cite{caricato2003parallel} reason about it on candidate solutions.

% There has been a a lot of work in motion planning for manipulators and movable objects. 
Navigation among movable objects deals with the combinatorial challenges of multiple objects~\cite{wilfong1991motion,van2009path} and has been shown to be a hard problem, and extended to manipulation applications~\cite{stilman2007manipulation}. Despite a lot of interesting work on challenges of manipulation and grasp planning, the current work shall make assumptions that avoid complexities arising from them. Manipulators opened the applications of rearrangement planning~\cite{ben1998practical,ota2004rearrangement}, including instances where objects can be grasped only once or monotone~\cite{stilman2007manipulation}, as well as non-monotone instances~\cite{havur2014geometric,srivastava2014combined}. Efficient solutions to assembly planning problems~\cite{Wilson:1994fk,Halperin:2000uq} typically assumes monotonicity, as without it the problem becomes much more difficult. Recent work has dealt with the hard instances of task planning~\cite{berenson2011task,cohen2014single} and rearrangement planning~\cite{krontiris2015dealing,krontiris2016efficiently,193}. Sampling-based task planning has made a recent push towards guarantees of optimality~\cite{vega2016asymptotically,schmitt2017optimal}. These are broader approaches that are invariant to the combinatorial structure of the application domain. The current work draws inspiration from these varied lines of research.

>>>>

%Related efforts:
%
%\begin{itemize}
%\item multi-robot planning and coordination
%\item manipulation task planning in general
%\item rearrangement 
%\item combinatorial literature
%\end{itemize}
The current work dealing with dual-arm object rearrangement touches upon the challenging intersection of a variety of rich bodies of prior work. It is closely related to multi-robot planning and coordination where a challenge is the high dimensionality of the configuration space. %Early work has suggested to slightly reduce the~\cite{aronov1999motion} tried dimensionality reduction strategies. 
Optimal strategies were developed for simpler instances of the problem~\cite{SolYuZamHal15}, although in general the problem is known to be computationally hard~\cite{solovey2016hardness}. Decentralized approaches~\cite{van2005prioritized} also used velocity tuning~\cite{leroy1999multiple}. General multi-robot planning tries to plan for multiple high-dimensional platforms \cite{wagner2012probabilistic,Gharbi:2009fu} using sampling-based techniques. 
Recent advances provide scalable~\cite{SoloveySH16:ijrr} and asymptotically optimal~\cite{Dobson:2017aa,Shome2019} sampling based frameworks. 
%Assembly planning \cite{halperin2000general,sundaram2001disassembly} is another related problem dealing with multi-body planning.

@@

%\textbf{The Traveling Salesman Problem}:
In this work we leverage a connection between dual-arm rearrangement and two combinatorial problems: (1) optimal matching~\cite{edmonds1965maximum} and (2) TSP. On the surface the problem seems closely related to multi-agent TSP.
%This can itself be tweaked with assumptions on same or different start and goal positions, symmetric or asymmetric metrics, directed or undirected graphs or posing the problem as an optimization
%A seminal paper~\cite{frederickson1976approximation} provides the formulation for k-TSP, with solutions that split a single tour.
%Prior work \cite{rathinam2006matroid} has formulated the problem of multi-start to multi-goal k-TSP as an optimization task. 
Prior work has formulated the k-TSP solution in terms of splitting a single tour~\cite{frederickson1976approximation} or as an optimization task~\cite{rathinam2006matroid}.
Some work~\cite{friggstad2013multiple} deals with asymmetric edge weights which are more relevant to the problems of our interest. 
%\textbf{Errand Scheduling}:
%If the object transfer is seen as an errand \cite{slavik1997errand} then a graph constructed with nodes being object-arm assignments and edges implying the order of coordinated execution. 
%There has been prior work \cite{garg2000polylogarithmic} that operates over the assignment graph(where pairwise assignments are nodes), an object's group consists of sets of nodes that include an object
The problem can be posed as an instance of multi vehicle pickup and delivery (PDP)~\cite{parragh2008survey}. 
%Most of the work has been motivated by ride-sharing applications and timed task scheduling, so there are a lot of variations that deal with time windows but this line of research provides some ILP formulations for the problem.
Prior work~\cite{coltin2014multi} has applied the PDP problem to robots, taking into account time windows and robot-robot transfers.
The seminal work~\cite{lenstra1981complexity,savelsbergh1995general} has explored its complexity, and concedes to the hardness of the problem, while others studied cost bounds~\cite{TrePavFra13}, and proposed ILP formulations~\cite{savelsbergh1995general}. 
%The existing work that delves into the combinatorial challenges of the problem, do not account for the costs incurred due to coordination of the agents during task execution. 
%Pickup and Delivery problems under track contention \cite{caricato2003parallel}, makes an attempt to reason about the coordination of the robots over solutions obtained with Tabu Search.
Typically this line of work ignores coordination costs, though some methods~\cite{caricato2003parallel} reason about it on candidate solutions.

% There has been a a lot of work in motion planning for manipulators and movable objects. 
Navigation among movable objects deals with the combinatorial challenges of multiple objects~\cite{wilfong1991motion,van2009path}, and has been shown to be a hard problem, and extended to manipulation applications~\cite{stilman2007manipulation}. Despite a lot of interesting work on challenges of manipulation and design choices~\cite{littlefield2016evaluating} in grasp planning, the current work shall make assumptions that avoid complexities arising from them. The availability of manipulators opened the door to solve rearrangement tasks~\cite{ben1998practical,ota2004rearrangement}, including instances where objects can be grasped only once or monotone~\cite{stilman2007manipulation}, as well as non-monotone instances~\cite{havur2014geometric,srivastava2014combined}. Efficient solutions to assembly planning problems~\cite{Wilson:1994fk,Halperin:2000uq} typically assume monotonicity, as without it the problem becomes much more difficult. Recent work has dealt with the hard instances of task planning~\cite{berenson2011task,cohen2014single} and rearrangement planning~\cite{krontiris2015dealing,krontiris2016efficiently,193}. Sampling-based task planning has made a recent push towards guarantees of optimality~\cite{vega2016asymptotically,schmitt2017optimal}. These are broader approaches that are invariant to the combinatorial structure of the application domain. The current work draws inspiration from these varied lines of research.
